### 2019-03-21 e5_imbalance.tex
More text for evaluation section.

\begin{figure*}[t]

\subfloat[Imbalance 0.5.\label{fig:efficiency-imbalance-0.5}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_0.5.pdf}
}
\subfloat[Imbalance 1.0.\label{fig:efficiency-imbalance-1.0}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.0.pdf}
}

\subfloat[Imbalance 1.5.\label{fig:efficiency-imbalance-1.5}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_1.5.pdf}
}
\subfloat[Imbalance 2.0.\label{fig:efficiency-imbalance-2.0}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}
}

\caption{Task granularity vs efficiency for varying load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Lower is better.\label{fig:efficiency-imbalance}}
\end{figure*}