### 2019-03-22 f5_task_graphs.tex
Add figures for task graphs.

\begin{figure}[t]

\subfloat[Trivial.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/trivial.pdf}
}
\hfill
\subfloat[Stencil.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/stencil.pdf}
}
\hfill
\subfloat[FFT.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/fft.pdf}
}

\hspace*{\fill}
\subfloat[Sweep.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/dom.pdf}
}
\hfill
\subfloat[Tree.]{
\includegraphics[width=0.45\columnwidth]{figs/sample-task-graphs/tree.pdf}
}
\hspace*{\fill}

\caption{Sample task graphs.\label{fig:task-graphs}}
\end{figure}