### 2019-03-26 e5_imbalance.tex
Cut down on the number of load imbalance plots.

\begin{figure}[t]

\centering
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}

\caption{Task granularity vs efficiency under load imbalance (nearest pattern, 5 dependencies per task, 4 task graphs, 1 node). Lower is better.\label{fig:efficiency-imbalance}}
\end{figure}