### 2019-05-24 reviews/response.md
revise response.md

We thank the reviewers for their detailed and insightful feedback.

We have added four new implementations to Task Bench: Dask, MPI bulk synchronous,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section, including preliminary results from experiments with GPUs on
the Piz Daint supercomputer.

Major changes in the revised paper:

  * New implementations: Dask, MPI-bulk sync, MPI+OpenMP and MPI+CUDA
  * New sections
      - Implementation: sections 4.3 (Dask) and 4.5 (MPI+X)
      - Evaluation: sections 5.6 (weak/strong scaling) and 5.9 (GPUs)
  * New figures
      - Listings 1 and 2 (code samples)
      - Figures 11, 12, 15 (experiments)

In response to specific reviewer questions:

Reviewer 1:

The memory bound kernel has been written to perform a varying number
of passes (or partial passes) over a fixed-sized buffer, in order to
keep working set size constant and avoid interference due to cache
effects, which in our opinion makes the results difficult to
interpret.

See the code samples in Listing 1 and 2.

Reviewer 2:

Figures 3 and 4 have been updated and new text added.

Reviewer 3:

Weak/strong scaling and TPS are all application and system dependent
metrics. METG differs from these metrics in that can simultaneously
(a) capture the behavior of the sytem at the limit of its scalability
(which is not true of weak/strong scaling) and (b) while achieving a
specified amount of useful work (which is not true of TPS). METG also
isolates the impact of changes to the communication topology at scale,
whereas in a strong scaling study this effect would be entangled with
issues of system overhead. Examples of how to compute METG from
application wall time have been added to Section 2.

METG captures the average task granularity that can be supported. It
is acceptable for some tasks to be smaller and others larger, as
overheads will even out across tasks.

A comparison with weak/strong scaling has been added in Section 5.6.

Reviewer 4:

We have implemented a MPI-bulk sync style by adding a MPI_Barrier after 
computation to make sure allcomputations are done before communication. 
Besides, we have also implemented a MPI+OpenMP,which also can be
considerred as BSP mode as all OpenMP threads are synchronized before
entering communication stage.

One of the salient features of Task Bench is that it permits the
implementation of a benchmark with M configurations and N systems with
an amount of effort that is O(M+N) rather than O(MN). We have updated
the introduction to reflect this.

The tasks in Task Bench are idealized compute or memory-bound loops;
problem size is simulated by controlling the number of iterations in
these loops. Listing 1 is an example.

Shared memory is permitted and is used in our MPI+OpenMP
implementation (among others).

The code for determining the dependence pattern and implementations of
kernels are contained in the core. A code sample from an
implementation is included in Listing 2. As the reviewer can see,
implementations can be quite simple.

TODO: explain MPI+CUDA?

Reviewer 5:

As described in Section 6, citations [27] and [37] are the only papers
we're aware of which provide an empirical performance evaluation of a
substantial number of systems. Many mini-app and benchmark papers only
include a small number of systems, or describe systems but do not
include an empirical evalution of performance. When there are
follow-up papers which describe additional implementations, they
rarely make comparisons beyond the reference implementation of the
code.

The output of a task is an array of bytes overwritten with a repeating
pattern (timestep, point). If the input does not match an assertion is
thrown.

Any measure of efficiency can be used with METG. For example, a
mesh-based application could use mesh cells processed per second (as a
percentage of the highest throughput achieved with any problem
size). We measured flops and memory bandwidth as 
peak performance with Task Bench as it gives us a
straightforward measure that provides assurance that all
implementations are configured correctly.