### 2019-07-23 t2_parameters.tex
Add table of task bench parameters.

\begin{table}[t]
\small

General options:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Values & Purpose \\
\hline

height & height of graph & number of timesteps \\
width & width of graph & degree of parallelism \\
dependence & trivial, stencil, etc. & dependence relation \\
kernel & compute, memory, etc. & type of kernel \\
output & bytes per dependency & degree of comm.
\end{tabular}

\vspace{1em}

Options specific to dependence relations:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Valid For & Purpose \\
\hline
radix & nearest & dependencies per task
\end{tabular}

\vspace{1em}

Options specific to kernels:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Valid For & Purpose \\
\hline
iterations & compute, memory & task duration/problem size \\
span & memory & bytes used per task per iter. \\
scratch & memory & total working set size \\
imbalance & load imbalanced & degree of imbalance
\end{tabular}

\caption{Task Bench parameters.\label{tab:parameters}}
\vspace{-0.5cm}
\end{table}