### 2019-07-25 0_abstract.tex
Changes to trim word count.

\begin{abstract}
% 148 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization allows a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

We conduct a comprehensive study with implementations of Task Bench in
15 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce and use a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each system
and the impact of different dependence patterns. We also
study each system's scalability, ability to hide communication and
mitigate load imbalance.
\end{abstract}