### 2019-07-30 0_abstract.tex
Tweak abstract to make performance claims.

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization enables a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

We conduct a comprehensive study with implementations of Task Bench in
15 programming systems on up to 256 Haswell nodes of the Cori
supercomputer. We introduce a novel metric, \emph{minimum effective task
  granularity} to study the baseline runtime overhead of each
system. We show that when running at scale, 100~\textmu{}s is the
smallest granularity that even the most efficient systems can reliably
support with current technologies. We also study each system's
scalability, ability to hide communication and mitigate load
imbalance.
\end{abstract}