### 2019-11-21 t2_parameters.tex

\begin{table}[t]
\small

General options:

\vspace{0.25em}

\begin{tabular}{l | l | l}
Parameter & Values & Purpose \\
\hline

height & height of graph & number of timesteps \\
width & width of graph & degree of parallelism \\
dependence & trivial, stencil, etc. & communication pattern \\
\quad \raisebox{0.35ex}{$\llcorner$} radix & \quad (for nearest pattern) & dependencies per task \\
kernel & compute, memory, etc. & type of kernel \\
\quad \raisebox{0.35ex}{$\llcorner$} iter. & \quad (for all kernels) & task duration \\
\quad \raisebox{0.35ex}{$\llcorner$} span & \quad (for memory kernel) & bytes used per task per iter. \\
\quad \raisebox{0.35ex}{$\llcorner$} scratch & \quad (for memory kernel) & total working set size \\
\quad \raisebox{0.35ex}{$\llcorner$} imbal. & \quad (for load imbalance) & degree of imbalance \\
output & bytes per dependency & degree of comm.
\end{tabular}

\caption{Task Bench parameters.\label{tab:parameters}}
\vspace{-0.75cm}
\end{table}