### 2020-02-04 pldi20_reviews/rebuttal.md
Edit rebuttal.

We thank the reviewers for their feedback.

Responding to reviewer A on novelty, we provide a quantification (via
METG) of typical programming system overheads. Contrary to the
suggestion of reviewer A, we found that the METGs we found in practice
did not always match our intuitions. In fact, some were off to such a
large degree (multiple orders of magnitude) that they prompted the
development teams to conduct major new development to address our
results (more on this below).

METG itself is also a contribution. Without METG, it is difficult to
quantify the programming system's contribution to application
overheads. Papers that look at system overheads typically report
measurements in tasks per second, but either use empty tasks or do not
report the level of efficiency achieved. (See below for citations.) In
either case it is impossible to determine how much useful work is
being performed. METG constrains efficiency so that a fixed amount of
work is being performed relative to overhead.

METG is parameterized on the efficiency metric. This is necessary, as
it would not make sense to measure FLOPS on a memory bound benchmark,
or vice versa. However, in spite of this, we find that the METG values
tend to fall into a certain characteristic range which is consistent
across efficiency measures. For example, compare Figures 6 and 8. The
systems in each graph occupy roughly the same parts of each graph, and
in most cases the relative ordering of systems is the same (with a
small number of close ties flipped between one and the other). This
holds when plotting Figure 8 on axes of efficiency vs task granularity
(which we will include in the final paper).

METG is also specific to the machine configuration: it can vary with
network, CPU, etc. However it is again our experience that METG for a
given programming system tends to span a range of typical values, and
that these can guide both application and system development. Also,
more precise measurements can identify specific bugs as noted below.

Beyond this, it would simply be impossible to conduct any quantitative
evaluation of a large number of systems without Task Bench. This is a
major contribution. Task Bench permits $\mathcal{O}(N+M)$ rather than
$\mathcal{O}(NM)$ implementation effort. In looking at the related
work, we find that the only benchmarks with a comparable number of
implementations are substantially more limited than Task Bench.

Reviewers A and D note that the bugs described in lines 197-199
currently do not have a larger analysis. We include a draft of this
analysis below and will incorporate it into the final paper.

Reviewer B asks about the impact of the network. It is our experience
that, except at the absolute lowest METGs, most programming system
overhead is on the CPU. Thus our study can highlight ways in which
these systems can be optimized to reduce unnecessary
overheads. However, we agree that comparing networks is a good
direction for future work.

TPS is reported in citations \[6, 28, 32]. See also \[A-C] below.

Fitting Task Bench configurations to specific applications is a
direction for future work, but the patterns in Figure 1 are derived
from common scientific workloads (and this is an extensible set).

Irregular parallelism is enabled by allowing dependencies to vary over
time. For example, Task Bench includes a randomized dependence
pattern.

An excerpt from the OpenMP implementation is included below. Note all
our implementations are open source and a link will be included in the
final paper.

Responding to reviewer C: the task graph is streamed whenever
possible. For example, both Realm and Dask operate on explicit task
graphs, but Realm is streamed (the API permits executing one chunk of
the graph while the next is being constructed), whereas Dask requires
the entire graph to be constructed. But see also our note on Dask
below.

Problem size is the `iterations` parameter in Listing 1. Note that
since these are idealized kernels, what matters is the total number of
FLOPs executed, which directly corresponds to `iterations`.

We use weak scaling. Height is fixed at 1000 (so the duration of the
run remains constant) while width varies with the number of
processors.

MPI p2p is shown in Listing 2.

## Analysis of Performance Bugs Found

Realm and Chapel use DMA subsystems that are optimized for large
copies. Early Task Bench experiments revealed high METG values that
were diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by 2x in the case of
Chapel. These improvements affect any application where fine-grained
tasks are used.

Further analysis of Realm efficiency indicated many overheads due to
dynamic graph construction. The Realm developers implemented a new
"subgraph" API in response to this feedback to amortize the cost of
repeatedly constructing isomorphic task graphs. This API is used in
Task Bench to achieve further speedups, up to nearly an order of
magnitude.

Analysis of the schedulers in Chapel and X10 revealed that certain
poor scheduling decisions could result in poor utilization of the CPUs
on a node. This resulted in poor peak performance, even at large task
granularities, which in some cases made it impossible to measure METG
(because peak efficiency did not exceed 50%). We reported these issues
to the systems' developers and used fixes or workarounds in our
experiments.

PaRSEC uses a task pruning algorithm to improve scalability at large
node counts. Initial Task Bench results achieved less than the
expected scalability: METG was rising too quickly with node
count. This was diagnosed by the PaRSEC developers as a bug in task
pruning, and subsequently fixed. We also implemented a version of the
PaRSEC code, `shard`, which uses manual pruning to demonstrate that
additional gains may be possible.

Early Task Bench results for Dask revealed that the cost of scheduling
a task was $\mathcal{O}(N)$ where $N$ is the number of tasks in a task
graph, causing overall cost for a task graph with $N$ nodes to be
$\mathcal{O}(N^2)$. This issue was reported to and confirmed by the
Dask developers. As a work around, our result in the paper use a
lower-level interface which does not suffer from this asymptotic
slowdown.

Initial experiments with TensorFlow revealed that the TensorFlow Task
Bench implementation was not able to use all cores on a node (thus
making it impossible to measure METG since the efficiency was below
50%). This was diagnosed by the TensorFlow developers as an issue in
constant folding. The entire Task Bench task graph was being detected
as constant, and the constant-folding pass runs on one core. As a
workaround, the developers suggested marking the task graph inputs as
non-constant so that TensorFlow's normal task scheduler could be used.

## OpenMP Code Excerpt

```c++
void issue_task(Graph g, long t, long p,
                char *output,
                char **inputs, long num_inputs,
                char *scratch) {
  switch (num_inputs) {
  case 1:
    {
      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(inout: scratch)
      g.execute_point(t, p, output, inputs, scratch);
    }
    break;
  case 2:
    {
      #pragma omp task depend(inout: output) depend(in: inputs[0]) depend(in: inputs[1]) depend(inout: scratch)
      g.execute_point(t, p, output, inputs, scratch);
    }
    break;
  // ...
  }
}

void execute_task_graph(Graph g) {
  char **outputs = (char **)malloc(g.width * sizeof(char *));
  char **scratch = (char **)malloc(g.width * sizeof(char *));
  char **inputs = (char **)malloc(g.width * sizeof(char *));
  char ***input_ptrs = (char ***)malloc(/* ... */);
  // initialize data structures...

  for (long t = 0; t < g.height; ++t) {
    for (long p = 0; p < g.width; ++p) {
      if (g.contains_point(t, p)) {
        long idx = 0;
        for (long dep : g.deps(t, p)) {
          input_ptrs[p][idx] = inputs[dep];
          idx++;
        }

        issue_task(g, t, p, outputs[p], input_ptrs[p], idx, scratch[p]);
      }
    }
    std::swap(inputs, outputs);
  }
}
```

## Additional Citations

\[A]: Mashayekhi, Omid, Hang Qu, Chinmayee Shah, and Philip
Levis. "Execution Templates: Caching Control Plane Decisions for
Strong Scaling of Data Analytics." In 2017 USENIX Annual Technical
Conference (USENIX ATC 17), pp. 513-526. 2017.

\[B]: Moritz, Philipp, Robert Nishihara, Stephanie Wang, Alexey
Tumanov, Richard Liaw, Eric Liang, Melih Elibol et al. "Ray: A
Distributed Framework for Emerging AI Applications." In 13th USENIX
Symposium on Operating Systems Design and Implementation (OSDI 18),
pp. 561-577. 2018.

\[C]: Sadooghi, Iman, Geet Kumar, Ke Wang, Dongfang Zhao, Tonglin Li,
and Ioan Raicu. "Albatross: An Efficient Cloud-Enabled Task Scheduling
and Execution Framework Using Distributed Message Queues." In 2016
IEEE 12th International Conference on e-Science (e-Science),
pp. 11-20. IEEE, 2016.