### 2020-03-23 e6_communication.tex

\begin{figure*}[t!]

\subfloat[16 bytes per task dependency.\label{fig:efficiency-communication-16}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_16.pdf}
}
%% \subfloat[256 bytes per task dependency.\label{fig:efficiency-communication-256}]{
%% \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_256.pdf}
%% }
%%
%% \subfloat[4096 bytes per task dependency.\label{fig:efficiency-communication-4096}]{
%% \includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_4096.pdf}
%% }
\subfloat[65536 bytes per task dependency.\label{fig:efficiency-communication-65536}]{
\includegraphics[width=\columnwidth]{figs/task-bench-results/communication/efficiency_nodes_64_comm_65536.pdf}
}

\vspace{-0.15cm}
\caption{Efficiency vs task granularity for varying communication (spread, 5 deps/task, 4 graphs, 64 nodes). Higher is better.\label{fig:efficiency-communication}}
\vspace{-0.05cm}
\end{figure*}