### 2020-05-27 e5_imbalance.tex
Mark additions.

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{figs/task-bench-results/imbalance/efficiency_imbalance_2.0.pdf}

\vspace{-0.15cm}
\caption{\color{blue} Efficiency vs task granularity under load imbalance (nearest pattern, 5 deps/task, 4 task graphs, 1 node). Higher is better.\label{fig:efficiency-imbalance}}
\vspace{-0.1cm}
\end{figure}