### 2020-05-27 sc20_reviews/response.md
Responses to reviewer 1.

We thank the reviewers for their detailed and insightful feedback.

Changes made in response to reviewer comments:

  * Added Section V.G: Load Imbalance.
  * Expanded Section VI: Case Study.

Responding to reviewer questions:

## Reviewer 1

  * Task graphs can be arbitrary, including irregular graphs. We added an example to Figure 1.

  * METG has already proven useful for exposing behaviors at the limits of system performance. We updated Section VI to clarify.

  * **FIXME: Please justify the use of small task granularity values (lower or equal than 1ms).**

## Reviewer 2

## Reviewer 3

## Reviewer 4

## Reviewer 5