### 2020-05-27 sc20_reviews/response.md
Update text on validation.

We thank the reviewers for their detailed and insightful feedback.

Changes made in response to reviewer comments:

  * Added Section V.G: Load Imbalance.
  * Expanded Section VI: Performance Issues Discovered.

Responding to reviewer questions:

## Reviewer 1

  * Task graphs can be arbitrary, including irregular graphs. We added an example to Figure 1.

  * METG has already proven useful for exposing behaviors at the limits of system performance. We updated Section VI to clarify.

  * **FIXME: Please justify the use of small task granularity values (lower or equal than 1ms).**

## Reviewer 2

## Reviewer 3

  * We added commit IDs to Table 5.

  * We renamed Section VI.

  * We added the Deakin citation to our related work.

## Reviewer 4

  * Text has been added to Section II to clarify validation.

## Reviewer 5