### 2020-06-30 e11_metg_predict_strong.tex

\begin{table}[t]
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std.~dev. & mean & std.~dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.24\times$ & 0.310 \\
spread  & $1.39\times$ & 0.322 & $1.30\times$ & 0.171 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.42\times$ & 0.723
\end{tabular}

\vspace{-0.20cm}
\caption{Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}