### 2020-06-30 f5_task_graphs.tex

\begin{figure}[t]

\subfloat[Trivial.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/trivial.pdf}
}
\hfill
\subfloat[Stencil.\label{fig:task-graphs-stencil}]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/stencil.pdf}
}
\hfill
\subfloat[FFT.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/fft.pdf}
}

{\captionsetup[subfloat]{farskip=-0.15cm,captionskip=-0.15cm}
\subfloat[Sweep.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/dom.pdf}
}
\hfill
\subfloat[Tree.]{
\includegraphics[width=0.32\columnwidth]{figs/sample-task-graphs/tree.pdf}
}
\hfill
\subfloat[Random.]{
\includegraphics[width=0.28\columnwidth]{figs/sample-task-graphs/random.pdf}
}
}

\vspace{-0.25cm}
\caption{Sample task graphs.\label{fig:task-graphs}}
\vspace{-0.2cm}
\end{figure}