### 2020-08-24 0_abstract.tex

% 149 words

\begin{abstract}
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of distributed programming systems under a
variety of application scenarios. Task Bench dramatically lowers the
barrier to benchmarking and comparing multiple programming systems by
making the implementation for a given system orthogonal to the
benchmarks themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore, Task Bench's
parameterization enables a wide variety of benchmark scenarios that
distill the key characteristics of larger applications.

To assess the effectiveness and overheads of the tested systems, we
introduce a novel metric, \emph{minimum effective task granularity}
(METG). We conduct a comprehensive study with 15 programming systems
on up to 256 Haswell nodes of the Cori supercomputer. Running at
scale, 100\textmu{}s-long tasks are the finest granularity that any
system runs efficiently with current technologies. We also study each
system's scalability, ability to hide communication and mitigate load
imbalance.
\end{abstract}