### 2020-08-27 e11_metg_predict_strong.tex
Update number of systems in validation comparison.

\begin{table}[t]
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std.~dev. & mean & std.~dev. \\
\hline
stencil & $1.64\times$ & 0.278 & $1.26\times$ & 0.246 \\
nearest & $1.75\times$ & 0.268 & $1.15\times$ & 0.331 \\
spread  & $1.43\times$ & 0.288 & $1.29\times$ & 0.206 \\
nearest, 4 graphs & $1.96\times$ & 0.680 & $1.26\times$ & 0.702
\end{tabular}

\vspace{-0.20cm}
\caption{Factor of separation between limit-ideal and limit-actual intersections across all 12 programming systems tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}